Object Categorization based on RGBD data

### 2014-02-04 Disseration/architecture.tex
aha

\chapter{Design}
\addcontentsline{toc}{chapter}{Design}

@@

	\subsubsection{Configurability}
	The programme is configurable through a text file of a standard *.ini format. It should be formatted as follows:
	\begin{tabular}{c}
	 param1 = 2
	 param2 = aha
	 [section]
	 param3 = no
	\end{tabular}

@@

>>>>

\chapter{Design and Implementation}
%\addcontentsline{toc}{chapter}{Design}

@@

	\subsubsection{Configurability}
	The programme is configurable through a text file of a standard *.ini format. It should be formatted as follows:\\ \ \\
	 param1 = 2\\
	 param2 = aha\\
	 {[section]}\\
	 param3 = no\\

@@

\section{Implementation}

	The complete program consists of more than 5600 lines of code in C++ and additional several hundreds of code in data preprocessing scripts.

	\subsection{Data preprocessing scripts}
		\begin{description}
			\item[imgListPrep.py] A script in python we have written for the purpose of listing hundreds of images (colour images and depth maps) and their corresponding labels in text files. These files are later used for batch processing such as training, but also for further data preprocessing.
			\item[b3doPrep] A short C++ programme used for object extraction from the B3DO dataset. It uses TinyXML, a lightweight C++ library for XML processing. It reads lists of files(depth maps, colour images and image annotations stored in the XML format) and iterates over them. Each iteration consists of loading an image, extracting labels and locations of bounding boxes from the XML, extracting the regions of interests and saving them at a new location.
			\item[divImgsByCats.py] Another short python script. Used mainly for splitting all the extracted objects into separate folders, one for every category.
			\item[pcdPrep] One more short C++ application. It was created for the reason of handling the Tokyo dataset. The dataset comes in a rather difficult-to-manage form. This application processes the raw files and saves them as point clouds in a PointCloud Library's PCD format.	 
			\item[grid.py and easy.py] These python scripts are an integral part of the libsvm library. Usually distributed as executables with the library, they help to perform corssvalidation and avoid overfitting. We have modified both scripts significantly in order to tune their performance on the very specific hardware describe in the Experiments section.
		\end{description}

		While the data prepared by pcdPrep is readily available for usage by the main programme the case with the B3DO dataset is quite different. The colour images and depth maps still have to be converted to point clouds because any further processing can be undertaken. That is because if the application were to work on-line with a Microsoft Kinect plugged in, it would have to handle the raw data from Kinect, which comes in a form of two images.

	\subsection{Conversion of colour images and depth maps into point clouds}

		The colour data and the depth data come from different cameras. The colour data come from 640 x 320 RGB camera, while the depth data come from active 320 x 240 infrared sensor. Only after registration the depth data is interpolated to match the colour data in resolution. Even after that operation, however, individual points from both images does not match exactly. That might introduce noise into the point cloud. We have not solved this issue and so that would be an area for further improvement. It did, however, focused our attention on another problem. If the two images do not fit perfectly it might be the case that one of the cameras have registered some points that had been invisible for the other one. Precisely, if there are any points in the colour image that the depth camera have not captured their depth would be meaningless. The Kinect's firmware does, in fact, address this issue. The depth value of every such point is set to NaN or Not a Number values. Before proceeding to the next stage we filter out all points, whose depth value has been set to NaN;

		Another problem is that only the Z coordinate is given explicitly in the depth map. The XY coordinates have to be inferred from the Z coordinate and the point's location relative to the sensor. We assumed that the sensor's position should correspond to the middle of the image or the the pixel at (320, 240). The Kinect specification says that the horizontal and vertical fields of view are respectively $57^{\circ}$ and $43^{\circ}$. We can compute the coordinates as: \\

		\begin{centering}
			$factorH = tg (\frac{57}{2}) * \frac{\pi}{180}$\\
			$factorV = tg (\frac{43}{2}) * \frac{\pi}{180}$\\
			$X = depth * factorH * (pixelX / 320)$\\
			$Y = depth * factorV * (pixelX / 240)$\\
			$Z = depth$\\
		\end{centering}

	\subsection{Input/Output Objects}

		As stated in the input/output section above, the application is very data dependent. Therefore, some objects for input/output operations were designed and implemented. The whole public API is delivered by the IoUtils class. It's methods allow reading and writing of text files, binary files of different formats, vectors of elements, vectors of vectors of elements and OpenCV's Mat objects. Elements that can be worked with are all C/C++ primitive types, all user defined types with defined std::string toString() method and OpenCV's Mat objects. In order to enhance performance, the IoUtils statically uses an object of type VectorIO for any vector input/output operation. The VectorIO is a template class parametrized by a type that is to be read into or written from a vector to a file on a hard drive. The VectorIO class uses a default implementation for all C/C++ primitive type and has template specialisations for the following types: std::vector, std::string, cv::Mat, any user defined object with a public std::string toString() method. Moreover, both IoUtils and VectorIO uses custom type trait classes that provide file extensions, size of types in bytes or processing algorithms for different data types.

### 2014-02-04 Disseration/experiments.tex
aha

\chapter{Experiments and Results}
\addcontentsline{toc}{chapter}{Experiments and results}

@@

	\subsection{Detectors}
	\rednote{run tests of SIFT, ISS, US}
	Keypoint detection algorithms have been compared with respect to repeatability, invariance \cite{pcl_keypoint_comparision} as well as object retrieval and recognition \cite{3d_keypoint_eval}. 3D-SIFT and ISS achieve the highest robustness for invariance. ISS appears to be the best choice for object retrieval. Dense detector (or uniform sampling) achieved the best performance in 2d image classification \cite{tsai2012bag}. Therefore, this paper compares SIFT, ISS and Uniform Sampling keypoint detection techniques.

	\subsection{Descriptors}
	\rednote{run tests on all PFH and SHOT descriptors and USC}
	PCL descriptors have been evaluated in terms of object retrieval performance \cite{pcl_features}. PFHRGB and SHOTCOLOR are the most successful. Simpler versions of those, that is: PFH, FPFH, SHOT and USC scored a little worse. We evaluate all of them in terms of BoW performance.

	\begin{table}[!ht]
	\centering	
	\includegraphics[width=.25\textwidth]{figs/desc_b3do}
	\caption{Highest accuracy obtained with FPFH, PFH and PFHRGB descriptors on the B3DO dataset}
	\label{tab:desc_b3do}
	\end{table}

	The figure \ref{tab:desc_b3do} shows highest accuracy obtained with the tested descriptors on the B3DO dataset. FPFH scored the best result, even tough it is an approximate and theoretically the least precise of all evaluated descriptors. On the other hand, the PFHRGB scored better than PFH. It does not surprise, since the colour information is believed to have discriminative value.

	\subsection{Codebook}
	\rednote{implement and run fisher}
	KMeans is used in almost every implementation of the Bag of Words image processing pipeline \cite{tsai2012bag, toldo2009bag}. Csurka \emph{et al} showed that the number of visual words has a significant impact on the final results \cite{csurka2004visual}. We run a number of tests to determinate an optimal codebook size.

@@

	The results from the figure \ref{fig:cluster_b3do} partially confirm Csurka's findings. The performance raises with the increasing number of centroids up to the point of 1500 centroids and 65.22\%. Then it starts to diminish. The discrepancy might be caused by the following factors: (1) ISS detector finds too many or irrelevant points, thus introducing noise or (2) The FPFH descriptor has too few dimensions (33) to be divided into more than 1500 regions in a meaningful way.

	\subsection{B3DO}
	Extraction of single objects yielded 78 separate categories with number of entries ranging from 1 (tape) to 299 (table). 8 categories were picked at random with a restriction that there should be at least 50 instances in each of them. Further, the objects were split into two sets (training and testing) with a proportion of 1:1. As names suggest, the estimation is performed on the test set and the evaluation on the test set.

	The highest accuracy achieved for this dataset is 64\%. Table \ref{tab:b3do_conf_matrix} contains a confusion matrix, number of examples per category and accuracy. The confusion matrix depicts misclassification errors. Let m\subscript{i, j} be an element of the confusion matrix at the i\superscript{th} row and j\superscript{th} column. It shows how many elements from the i\superscript{th} category was assigned to the j\superscript{th} category. A high value of m\subscript{i, j} such that $i \neq j$ indicates that the classifier cannot distinguish between those two classes.

@@

>>>>

\chapter{Experiments and Results}
%\addcontentsline{toc}{chapter}{Experiments and results}

@@

	\subsection{Detectors and descriptors}
	Keypoint detection algorithms have been compared with respect to repeatability, invariance in \cite{pcl_keypoint_comparision} and in terms of object retrieval and recognition in \cite{3d_keypoint_eval}. 3D-SIFT and ISS achieve the highest robustness for invariance, while the latter appears to be the best choice for object retrieval. Dense detector (or uniform sampling) achieved the best performance in 2D image classification \cite{tsai2012bag}. Therefore, this paper compares SIFT, ISS and Uniform Sampling keypoint detection techniques.

	PCL descriptors have been evaluated in terms of object retrieval performance in \cite{pcl_features}. PFHRGB and SHOTCOLOR delivered the best performance, while their simpler counterparts (PFH, FPFH, SHOT) were only a little worse. USC, which is a global descriptor and thus not suitable for the Bag of Words approach, achieved very similar accuracy. We focus on the family of PFH algorithms.

	\begin{table}[!ht]
	\centering	
	\includegraphics[width=.55\textwidth]{figs/desc_b3do}
	\caption{Highest accuracy obtained with FPFH, PFH and PFHRGB descriptors on the B3DO dataset}
	\label{tab:desc_b3do}
	\end{table}

	The best results achieved are shown in the figure \ref{tab:desc_b3do}. They were scored on the B3DO dataset. As for the detection algorithms, the ISS provided consistently the highest accuracy. The SIFT and Uniform Sampling provide very similar accuracy, but the latter is significantly faster. SIFT is in fact slowest of the three. As for the descriptors, the FPFH delivered the best performance even though it is an approximate and theoretically the least precise algorithm. It is also the fastest of the three. It appears that the combination of ISS and FPFH is not only surprisingly good in terms of accuracy, but it is also the fastest. These two algorithms allow near real time point-cloud processing and classification. PFHRGB and SIFT are several times slower.

	\subsection{Codebook}
	KMeans is used in almost every implementation of the Bag of Words image processing pipeline \cite{tsai2012bag, toldo2009bag}. Csurka \emph{et al} showed that the number of visual words has a significant impact on the final results \cite{csurka2004visual}. We run a number of tests to discover an optimal codebook size for our setting.

@@

	The results from the figure \ref{fig:cluster_b3do} partially confirm Csurka's findings. The performance raises with the increasing number of centroids up to the point of 1500 centroids and 65.22\%. The discrepancies might be caused by the following factors: (1) ISS detector finds too many or irrelevant points, thus introducing noise or (2) The FPFH descriptor has too few dimensions (33) to be divided into more than 1500 regions in a meaningful way.

	\subsection{B3DO}
	The majority of the point clouds in the B3DO dataset contain several different objects. We have extracted every individual object from the clouds and divided them into 78 separate categories. Unfortunately, they were badly balanced \emph{i.e.} the number of objects in a category varies from 1 (tape) to 299 (table). We have picked 8 categories at random with a restriction that there should be at least 50 object instances in a category. Then, the objects were split into two sets (training and testing) with a proportion of 1:1. As the names suggest, they are used for estimation and evaluation respectively.

	The highest accuracy achieved for this dataset is 64\%. Table \ref{tab:b3do_conf_matrix} contains a confusion matrix, number of examples per category and accuracy. The confusion matrix depicts classification errors. Let m\subscript{i, j} be an element of the confusion matrix at the i\superscript{th} row and j\superscript{th} column. It shows how many elements from the i\superscript{th} category was assigned to the j\superscript{th} category. A high value of m\subscript{i, j} such that $i \neq j$ indicates that the classifier cannot distinguish between those two classes.

@@

	In this paper we have tackled the issue of three dimensional point cloud classification with a bag of words-based approach. The main focus was put on the design and efficient implementation of the pipeline in a way that allows easy addition of algorithms and switching them at run-time. What is more, we have gone to great lengths in order to find scientific databases that are suitable for the task of point cloud classification. We believe that the presence of such databases is fundamental, for it allows meaningful comparison with other researchers. We have experienced the following problems: Firstly, point cloud processing is computationally expensive and inefficient. Due to the lack of structure in point clouds (\textit{i.e.} no predefined relationship between points) additional calculations have to be performed (\textit{e.g.} finding nearest neighbours requires distance comparison for all points in a cloud, whereas for a 2D image one instantly knows where adjacent pixels lie). Secondly, there are virtually no databases that would allow extensive training of the bag of words-based pipeline. Machine learning algorithms have ravenous appetite for labelled data. As the newest findings in the state-of-the-art 2D object classification suggest, an order of tens of thousands of samples is a prerequisite for any satisfactory results We had to make do with as few as 330 objects in our training set. Moreover, there is very little interest in the scientific community for this particular task. The majority of research is conducted on retrieval of CAD-like objects from mesh databases. Last, but not least, there is no 3D detection or description algorithm designed solely for the purpose of bag of words object classification.

	In conclusion I would not recommend any bag-of-words based approach classification of real point clouds registered with a low quality camera such as Microsoft Kinect. Being computationally expensive and rather inefficient, it delivers inferior performance in terms of accuracy and speed when compared to 2D RGB image based algorithms. To makes matters worse, generalisation of 2D algorithms to the 3D domain can be very involved, causing the cutting-edge inventions of the 2D image processing scientists unavailable for the purpose of 3D point cloud processing.

### 2014-02-04 Disseration/introduction.tex
aha

\chapter{Introduction}
\addcontentsline{toc}{chapter}{Introduction}

@@

	This paper tackles two issues. Firstly, the Bag of Words (or BoW for short) approach to image classification is reviewed. The Bag of Words method has originated from the natural language processing domain \rednote{ref}. It makes a strong assumption that the word occurrence in a text document define the meaning of the document. It were Csurka \textit{et al} who have first used BoW for image classification in 2004 \cite{csurka2004visual}. Since we cannot simply translate an image into natural language a pipeline that help us achieve this goal is needed. It usually consist of: keypoint detection, feature extraction, vector quantization and classification, steps that will be discussed later.

@@

	\section{scope and purpose of this work}
		The goal of this work is to design a flexible and configurable application for point cloud classification. The point cloud should be registered by the Microsoft Kinect RGBD camera. The scope of the work is as follows:
		\begin{itemize}
		\item literature review in the area of bag of words image classification
		\item design of a pipeline for bag of words classification of point clouds
		\item design of a configurable programme for point cloud classification
		\item implementation of the programme in C++ using OpenCV and/or PointCloud Library
		\item evaluation of algorithms for the bag of words classification
		\item evaluation of the programme on two scientific RGBD object databases
		\end{itemize}

\section{Bow of Words}

	The Bag of Words model is a an intermediate representation used in natural language processing, information retrieval and data mining \rednote{pub}. The model can be seen as a simplification, for it obliterates any grammatical information and even word order contained within an original text document. What is left is an unstructured group of words, usually in a form of a vector or a histogram of word incidence.

@@

	In natural language processing BoW is used to infer semantic meaning of documents \rednote{pub}. If we could translate an image into a text document we might be able to employ similar methods. A question arises: How does one make a text document from an image?

@@

		\rednote{describe fundamentals of keypoint detection in greater detail}
		The most common detectors make use of a Harris corner detector or image's first or second derivatives. A Harris-Laplace detector is an example of a Harris-based detector. The Harris function is scale adapted and its outcome is a subject to a Laplacian-of-Gaussian (LoG) operator, which selects relevant points in scale space. Images' regions' 2\textsuperscript{nd} derivatives, namely the regions' Hessians, can be combined with a LoG operator as well. This combination allows selection of points significant in the two spaces: the scale space and the Hessian's determinant space. The latter entails the speed at which pixel intensities change in the neighbourhood of a point.

@@

		Keypoint description takes form of coordinates in a high-dimensional space. One of the most precise and repeatable algorithm is SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 128-dimensional vector of floating point values. It is the most often extracted descriptor in BoW pipelines \rednote{ref}. Other methods include various colour descriptors, binary descriptors such as 512-dimensional GIST \cite{ponce2011cv}. There are techniques designed for 3D exclusively. Among them are Persistent Point Feature Histogram (PFH) \cite{pfh_rusu2008}, its faster alternative FPFH \cite{fpfh_rusu2009} and PFHRGB, which takes into account colour information, all implemented in PCL.

@@

		\subsubsection{Scene and Object Classification}
		Scene and object classification are among the most popular issues of computer vision nowadays \rednote{ref}. We would like to perform those tasks automatically for the sake of surveillance, navigation or automatic image tagging. The very first attempt to use BoW in scene classification was made by Csurka \textit{et al} in 2004 \cite{csurka2004visual}. They were inspired by an idea of a ``texton'', a building block of texture introduced earlier in pattern recognition and texture classification. The authors suggested a BoW pipeline described above. Then, they fed the resulting BoW histograms into two classifiers: Support Vector Machine (SVM) and Na\`ve Bayes (both discussed below). Fei-Fei \emph{et al} refined the original approach by examining several keypoint detectors and descriptors \cite{fei2005bayesian}. Moreover, they developed a novel probabilistic graphical model for classification. They achieved accuracy of 76\% on a large 13 category dataset.

@@

	\subsection{libsvm}
	\rednote{describe}
	\subsection{Apache log4cxx}
	\rednote{describe}

>>>>

\chapter{Introduction}
%\addcontentsline{toc}{chapter}{Introduction}

@@

	This paper tackles two issues. Firstly, the Bag of Words (or BoW for short) approach to image classification is reviewed. The Bag of Words method has originated from the natural language processing domain \cite{tsai2012bag}. It makes a strong assumption that the word occurrence in a text document define the meaning of the document. It were Csurka \textit{et al} who have first used BoW for image classification in 2004 \cite{csurka2004visual}. Since we cannot simply translate an image into natural language a pipeline that help us achieve this goal is needed. It usually consist of: keypoint detection, feature extraction, vector quantization and classification, steps that will be discussed later.

@@

	\section{Scope and purpose of this work}
		The goal of this work is to design a flexible and configurable application for point cloud classification. The point cloud should be registered by the Microsoft Kinect RGBD camera. The scope of the work is as follows:
		\begin{itemize}
		\item literature review in the area of bag of words image classification
		\item design of a pipeline for bag of words classification of point clouds
		\item design of a configurable programme for point cloud classification
		\item implementation of the programme in C++ using OpenCV and/or PointCloud Library
		\item evaluation of algorithms for the bag of words classification
		\item evaluation of the programme on two scientific RGBD object databases
		\end{itemize}

\section{Bag of Words}

	The Bag of Words model is a an intermediate representation used in natural language processing, information retrieval and data mining. The model can be seen as a simplification, for it obliterates any grammatical information and even word order contained within an original text document. What is left is an unstructured group of words, usually in a form of a vector or a histogram of word incidence.

@@

	In natural language processing BoW is used to infer semantic meaning of documents. If we could translate an image into a text document we might be able to employ similar methods. A question arises: How does one make a text document from an image?

@@

		The most common detectors make use of a Harris corner detector or image's first or second derivatives. A Harris-Laplace detector is an example of a Harris-based detector. The Harris function is scale adapted and its outcome is a subject to a Laplacian-of-Gaussian (LoG) operator, which selects relevant points in scale space. Images' regions' 2\textsuperscript{nd} derivatives, namely the regions' Hessians, can be combined with a LoG operator as well. This combination allows selection of points significant in the two spaces: the scale space and the Hessian's determinant space. The latter entails the speed at which pixel intensities change in the neighbourhood of a point.

@@

		Keypoint description takes form of coordinates in a high-dimensional space. One of the most precise and repeatable algorithm is SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 128-dimensional vector of floating point values. It is the most often extracted descriptor in BoW pipelines. Other methods include various colour descriptors, binary descriptors such as 512-dimensional GIST \cite{ponce2011cv}. There are techniques designed for 3D exclusively. Among them are Persistent Point Feature Histogram (PFH) \cite{pfh_rusu2008}, its faster alternative FPFH \cite{fpfh_rusu2009} and PFHRGB, which takes into account colour information, all implemented in PCL.

@@

		\subsubsection{Scene and Object Classification}
		Scene and object classification are among the most popular issues of computer vision nowadays. We would like to perform those tasks automatically for the sake of surveillance, navigation or automatic image tagging. The very first attempt to use BoW in scene classification was made by Csurka \textit{et al} in 2004 \cite{csurka2004visual}. They were inspired by an idea of a ``texton'', a building block of texture introduced earlier in pattern recognition and texture classification. The authors suggested a BoW pipeline described above. Then, they fed the resulting BoW histograms into two classifiers: Support Vector Machine (SVM) and Na\`ve Bayes (both discussed below). Fei-Fei \emph{et al} refined the original approach by examining several keypoint detectors and descriptors \cite{fei2005bayesian}. Moreover, they developed a novel probabilistic graphical model for classification. They achieved accuracy of 76\% on a large 13 category dataset.

@@

	\subsection{Apache log4cxx}
	Apache log4cxx is an efficient and easy to use logger for C++ programmes. It is based on a Apache log4j logger for Java applications. It provides a set of static methods for logging. It support simultaneous output to multiple destinations and have very well defined logging levels. It is perfectly documented as well.
